What We Eval in the Shadows

### 2021-04-05 paper/main.tex
Add few todos

The R language was intended to be extensible, the combination of lazy
evaluation, \c{substitute} and \c{eval} are the tools given to
developers to this end. This API is slightly more complex than just
passing a string, it is conceivable that this may discourage some
casual users. \Eval is also being used to reduce boilerplate code and
provide convenience features for programmers. We now give some
representative examples of its usage.

\mypara{Intercession.} A common use case for \eval is to be combined with
\c{match.call}. \c{match.call} walks up the call stack, captures the code that
invoked the currently executing function, and returns it as an unevaluated
expression. The pattern is to transform a call to some function \c{f} into a
call to \c{g} with some arguments retained and others modified. As an
illustration, consider the \c{vcpart} package's function \c{tvcglm} that is
translated to a call to \c{tvcm} with two modifications to the argument list:
argument \c{control} can't be missing and \c{fit} is set \c{"glm"}. The function
ends with a call to \c{eval.parent} to ensure that the rewritten call is
evaluted in the same environment the original call was.
\begin{lstlisting}
 tvcglm <- function(formula, data, family, ...) {
   k <- match.call()
   k[[1]] <- as.name("tvcm")
   if ("fit" %in% names(list(...))) warning("'fit' ignored.")
   k$fit <- "glm"
   eval.parent(k)
 }
\end{lstlisting}
This pattern is recognizable by the fact that the expression is a call
and the target environment is that of the parent.

\mypara{Code Generation.} The more traditional use of \eval is to execute code
that was assembled by the programmer into a string. Here we show the method
\c{plot} for class \c{sback} in package \c{wsbackfit}, simplified for
explanatory purposes. The function takes a long argument list, the names of
which are captured in the list \c{opt}. The string \c{stub} is composed of a
subset of the arguments passed to this function; the variable is used to
construct a call to \c{base::plot} which will draw a plot. \NOTE{This example is
  a bit messed up. What is var? Does the use of ... in the eval grab the
  arguments of this call? If yes does it not end up with xlab twice? HELP} The
\parse and \eval combination is used by the the \source function in the base
library to load R code from a file in the current workspace.
\begin{lstlisting}
 plot.sback <- function(x,...) {
   opt <- names(list(...))
   stub <- paste(
    ifelse("xlab" %in% opt,"",paste(",xlab=\"",var,"\"",sep="")),
    ifelse("type" %in% opt,"",",type=\"l\""), sep="")
   plot <- paste("plot(x.data,",stub,",...)",sep="")
   eval(parse(text=plot))
\end{lstlisting}
This pattern is recognizable by its use of \c{parse} to turn a string
into an expression.

\mypara{Debloating.} \Eval is often used as a means to reduce boilerplate code;
simple and repetitive code can easily be replaced with judcious use of \eval.
For example, the \c{data.table} package uses \eval to calls the \c{options}
function with named arguments taken from a vector of strings. While the benefits
are limited in this example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ...many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.

\mypara{Trivial.} When values are passed to \eval, they are returned
unchanged. They are an example of trivial uses of \eval. Another
trivial use is the empty expression, often found in JavaScript, but
rare in R.

>>>>

The R language was intended to be extensible. The combination of lazy
evaluation, \c{substitute} and \c{eval} are the tools given to
developers to this end. This API is slightly more complex than just
passing a string, it is conceivable that this may discourage some
casual users. \Eval is also being used to reduce boilerplate code and
provide convenience features for programmers. We now give some
representative examples of its usage.

\mypara{Intercession.} \todo{Pierre: update wrt the previous section if we want
to keep it. Also perhaps we could start with a simpler pattern and then show
this one} A common use case for \eval is to be combined with \c{match.call}.
\c{match.call} walks up the call stack, captures the code that invoked the
currently executing function, and returns it as an unevaluated expression. The
pattern is to transform a call to some function \c{f} into a call to \c{g} with
some arguments retained and others modified.  \todo{or change the target
function. Both are described as "patterns" in the R documentation.} As an
illustration, consider the \c{vcpart} package's function \c{tvcglm} that is
translated to a call to \c{tvcm} with two modifications to the argument list:
argument \c{control} can't be missing and \c{fit} is set \c{"glm"}. The
function ends with a call to \c{eval.parent} to ensure that the rewritten call
is evaluted in the same environment the original call was.  \begin{lstlisting}
tvcglm <- function(formula, data, family, ...) { k <- match.call() k[[1]] <-
as.name("tvcm") if ("fit" %in% names(list(...))) warning("'fit' ignored.")
k$fit <- "glm" eval.parent(k) } \end{lstlisting} This pattern is recognizable
by the fact that the expression is a call and the target environment is that of
the parent.

\mypara{Code Generation.} The more traditional use of \eval is to execute code
that was assembled by the programmer into a string. Here we show the method
\c{plot} for class \c{sback} in package \c{wsbackfit}, simplified for
explanatory purposes. The function takes a long argument list, the names of
which are captured in the list \c{opt}. The string \c{stub} is composed of a
subset of the arguments passed to this function; the variable is used to
construct a call to \c{base::plot} which will draw a plot. \NOTE{This example
  is a bit messed up. What is var? Does the use of ... in the eval grab the
arguments of this call? If yes does it not end up with xlab twice? HELP}
\todo{Does this comment still applies?} The \parse and \eval combination is
used by the the \source function in the base library to load R code from a file
in the current workspace.  \begin{lstlisting} plot.sback <- function(x,...) {
opt <- names(list(...)) stub <- paste( ifelse("xlab" %in%
opt,"",paste(",xlab=\"",var,"\"",sep="")), ifelse("type" %in%
  opt,"",",type=\"l\""), sep="") plot <-
  paste("plot(x.data,",stub,",...)",sep="") eval(parse(text=plot))
  \end{lstlisting} This pattern is recognizable by its use of \c{parse} to turn
  a string into an expression.

\mypara{Debloating.} \todo{Wouldn't reducing boilerplate be a better name?} \Eval is often used as a means to reduce boilerplate code;
simple and repetitive code can easily be replaced with judcious use of \eval.
For example, the \c{data.table} package uses \eval to calls the \c{options}
function with named arguments taken from a vector of strings. While the benefits
are limited in this example, it is an attractive tool for programmers.
\begin{lstlisting}
  opts = c("datatable.verbose"="FALSE", # ...many others
  for (i in names(opts))
    eval(parse(text=paste0("options(",i,"=",opts[i],")")))
\end{lstlisting}
This pattern is a special case of code generation, recognizable by the
fact that \eval is executed in a loop.

\mypara{Trivial.} When values are passed to \eval, they are returned
unchanged. They are an example of trivial uses of \eval. Another
trivial use is the empty expression, often found in JavaScript, but
rare in R. \todo{Given that we do not have any data I would get rid of it}
